Contextuality and the Born rule in fuzzy quantum theories

### 2017-04-06 prop_letter.tex
prop_letter.tex Here and there...

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\cite{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where any
subset of $\events$ is called a sub-event space if it satisfies the
follow conditions. 
\begin{itemize}
\item $\mathbb{0}\in\events'$. 
\item $\mathbb{1}\in\events'$. 
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}
Quantum IVPM not only stems from classical IVPM but also generalizes
quantum real-valued probability measures. Consider the collection
of intervals are all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
and a function~$\iota_{\infty}:\left[0,1\right]\rightarrow\mathscr{I}_{\infty}$
defined by $\iota_{\infty}\left(x\right)=\left[x,x\right]$. Then,
given a quantum real-valued probability measure~$\mu:\events\rightarrow\left[0,1\right]$,
$\bar{\mu}(P)=\iota\left(\mu(P)\right)$ is a quantum IVPM, and vice
versa. In another word, quantum IVPMs with $\mathscr{I}_{\infty}$
are just the same as quantum real-valued probability measures where
Gleason's theorem can apply.

@@

Recall when providing infinity measurement resource, the quantum probability
measures are simply induced by the Born rule for $d\ge3$. If we increase
the measurement resource, i.e., pick finer intervals, it is natural
to ask whether quantum IVPMs approach the Born rule. In order to understand
this question, it is instructive to consider the 2-dimensional non-Born
quantum probability measure~$\mu:\events\rightarrow[0,1]$ defined
as follow:
\begin{equation}
\mu_{0}(P)=\begin{cases}
\mu_{\ket{\ps}}^{B}(P) & \textrm{, if }P\in\events_{0}\textrm{ ;}\\
\mu_{\ket{0}}^{B}(P) & \textrm{, otherwis,}
\end{cases}
\end{equation}
where $\events_{0}=\left\{ \mathbb{0},\proj{\ps},\proj{\ms},\mathbb{1}\right\} $
is a sub-event space. In another word, $\mu_{0}$ is consistent with
the Born rule on every proper sub-event space, but not on the set
of all projectors.

Given a quantum IVPM~$\bar{\mu}$, the preceding idea motivates the
definition of \emph{core} relative to a sub-event space~$\events'$
as follow.
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in\events'.~\mu_{\rho}^{B}\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
Although the elements of a core are density matrices, we may just
use the state vectors for pure states if it is clear in the context.
For example, $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events_{0}\right)=\left\{ \ket{\ps}\right\} $,
where $\circ$ denote function composition.\newpage{}

@@

While the above example alerts that there might be a bad choice to
represent our observations as a quantum IVPM, this phenomena is purely
quantum, because Shapley proved that a good enough classical IVPM
always has a non-empty core.

\begin{thm}[Shapley~\cite{Shapley1971,Grabisch2016}]\label{thm:Shapley}Every
classical convex interval-valued probability measure has a non-empty
core, where $\bar{\mu}$ is called convex if 
\begin{equation}
\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\label{eq:QuantumInterval-valuedProbability-Convex}
\end{equation}
for all commuting $P_{0},P_{1}\in\events$, and the \emph{disjunction}~$P_{0}\vee P_{1}$
is defined to be $P_{0}+P_{1}-P_{0}P_{1}$~\cite{Griffiths2003}.\end{thm}

Since any classical IVPM we found has a non-empty core, we conjecture
that every classical IVPM has a non-empty core.

>>>>

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\cite{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where any
subset of $\events$ is called a sub-event space if it satisfies the
follow conditions. 
\begin{itemize}
\item $\mathbb{0}\in\events'$. 
\item $\mathbb{1}\in\events'$. 
\item For any projection~$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}
Quantum IVPM not only stems from classical IVPM but also generalizes
quantum real-valued probability measures. Consider the collection
of intervals are all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
and a function~$\iota_{\infty}:\left[0,1\right]\rightarrow\mathscr{I}_{\infty}$
defined by $\iota_{\infty}\left(x\right)=\left[x,x\right]$. Then,
given a quantum real-valued probability measure~$\mu:\events\rightarrow\left[0,1\right]$,
$\bar{\mu}(P)=\iota_{\infty}\left(\mu(P)\right)$ is a quantum IVPM,
and vice versa. In another word, quantum IVPMs with $\mathscr{I}_{\infty}$
are just the same as quantum real-valued probability measures where
Gleason's theorem can apply.

@@

Recall when providing infinity measurement resource, the quantum probability
measures are simply induced by the Born rule for $d\ge3$. If we increase
the measurement resource, i.e., pick finer intervals, it is natural
to ask whether quantum IVPMs approach the Born rule. In order to understand
this question, it is instructive to consider the 2-dimensional non-Born
quantum probability measure~$\mu_{0}:\events\rightarrow[0,1]$ defined
as follow:
\begin{equation}
\mu_{0}(P)=\begin{cases}
\mu_{\ket{\ps}}^{B}(P) & \textrm{, if }P\in\events_{0}\textrm{ ;}\\
\mu_{\ket{0}}^{B}(P) & \textrm{, otherwis,}
\end{cases}
\end{equation}
where $\events_{0}=\left\{ \mathbb{0},\proj{\ps},\proj{\ms},\mathbb{1}\right\} $
is a sub-event space. In another word, $\mu_{0}$ is consistent with
the Born rule on every proper sub-event space, but not on the set
of all projectors.

Given a quantum IVPM~$\bar{\mu}$, the preceding idea motivates the
definition of \emph{core} relative to a sub-event space~$\events'\subseteq\events$
as follow.
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in\events'.~\mu_{\rho}^{B}\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
Although the elements of a core are density matrices, we may just
use the state vectors for pure states if it is clear in the context.
For example, $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events_{0}\right)=\left\{ \ket{\ps}\right\} $
and $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events\right)=\emptyset$,
where $\circ$ denote function composition.

Notice that $\iota_{\infty}\circ\mu_{0}$ has a non-empty core relative
to $\events_{0}$ where every projectors are mutually commutative.
Recall commutative sub-event space reduces quantum IVPMs to classical
IVPMs. In general, we haven't found any classical IVPM has a non-empty
core, and Shapley has proved that a good enough classical IVPM always
has a non-empty core.

\begin{thm}[Shapley~\cite{Shapley1971,Grabisch2016}]\label{thm:Shapley}For
every interval-valued probability measure~$\bar{\mu}:\events'\rightarrow\mathscr{I}$,
if the following conditions are true.
\begin{itemize}
\item $\events'$ is commutative, i.e., $\bar{\mu}$ is classical.
\item For all commuting $P_{0},P_{1}\in\events$, we have $\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$,
where $P_{0}\vee P_{1}$ is defined to be $P_{0}+P_{1}-P_{0}P_{1}$~\cite{Griffiths2003}.
\end{itemize}
Then, $\mathrm{core}\left(\bar{\mu},\events'\right)\ne\emptyset$.\end{thm}

\newpage{}

@@
